Verification of Stencil Computations through Spatial Specifications

### 2016-07-03 paper.tex

>>>>

\subsection{Related work}

Various prior approaches to specifying and verifying the behaviour
of stencils have been based on fine-grained representations of the
code. These essentially duplicate the indexing scheme of a stencil
computation in their specifications.

\citet{kamil2016verified} propose a technique they call
\emph{verified lifting} for mapping low-level, possibly
greatly optimised, stencil computations into a high-level
predicate language using inductive program synthesis. 
The aim is that these high-level abstract specifications can
then be translated into high-performance DSLs, easing the
use of high-performance DSLs. Compared to our approach,
their specification language is more fine grained, capturing
the exact indexing pattern of a stencil. For example, the following
is an example post-condition in their language, generated by their
compiler \textsc{stng} from a Fortran stencil computation~\cite [p.3]{kamil2016verified}:
%
\begin{align*}
& \textit{post}(\texttt{a}, \texttt{b}) \equiv \forall \texttt{imin}+1 
\leq \texttt{i} \leq \texttt{imax}, \texttt{jmin} \leq \texttt{j} \leq
\texttt{jmax} \\
& \qquad \texttt{a}(\texttt{i},\texttt{j}) = 
\texttt{b}(\texttt{i}-1,\texttt{j}) + \texttt{b}(\texttt{i},\texttt{j})
\end{align*}
%
Noticeable, the bounds of the stencil and the exact indexing pattern
are captured. Our approach aims to be much more abstract, to
facilitate human reasoning, and avoid low-level lexical
errors. However, \textsf{stng} specifications are generated, \ie{}, 
are never written by a programmer, avoiding this
problem. Their inductive program synthesis approach allows
specifications to extracted even from optimised and unrolled code. For
example, the above post condition is generated from the code~\cite [p.3]{kamil2016verified}:
%
\begin{ExmVerbatim}
do j=jmin, jmax 
    t = b(imin, j)
    do i=imin+1,imax
        q = b(i, j)
        a(i, j) = q + t
        t = q
    end do
end do
\end{ExmVerbatim}
%
Thus, \textsc{stng} can understand cross-loop data-dependencies
in order to assign the above post-condition. CamFort is not currently
able to ascribe a specification in the above case as the use of
\texttt{t} on line 5 maps only to the definition on line 2, and no on
line 6. Extending our approach to follow loop-carried dependencies is
future work. An alternate approach for us would be to build ontop of
the work of Kamil~\emph{et al.} by inferring our specifications from 
\textsc{stng} (as an intermediate step).

In an alternate approach,~\citet{abe2013model}

There are many DSLs for stencil comptuation aimed at either
performance or verification, or both. Many elimiante out-of-bounds
errors, \eg{},~\cite{DBLP:journals/corr/abs-1109-0777}.
